---
title: "p8105_hw2_jdv2118"
output: github_document
---
```{r}
library(tidyverse)

library(readxl)

library(dplyr)

library(janitor)

library(tidyr)
```

# Problem 2

```{r}
mr_trash_wheel = 
  read_excel("data/Trash Wheel Collection Data.xlsx", "Mr. Trash Wheel", range = "A2:N550") %>%
  clean_names(dat = .) %>%
  mutate(sports_balls = 
  as.integer(
  round(sports_balls)), 
  trash_wheel = 1, 
  year =
  as.integer(year)) %>%
  drop_na(data = ., dumpster) %>%
  select(trash_wheel, everything())

head(mr_trash_wheel)

view(mr_trash_wheel)
```

```{r}
prof_trash_wheel = 
  read_excel("data/Trash Wheel Collection Data.xlsx", "Professor Trash Wheel", range = "A2:M97") %>%
  clean_names(dat = .) %>%
  mutate(trash_wheel = 2) %>%
  drop_na(data = ., dumpster) %>%
  select(trash_wheel, everything())

head(prof_trash_wheel)

view(prof_trash_wheel)
```

```{r}
trash_wheel_tidy =
  bind_rows(mr_trash_wheel, prof_trash_wheel)

view(trash_wheel_tidy)
```

# Problem 3 

```{r}
national_politicians =
  read_csv("data/pols-month.csv") %>%
  clean_names(dat = .) %>%
  separate(col = mon,   into = c("year",      "month", "day"), sep = "-") %>%
  mutate(
    month = recode(month, `01` = "January", `02` = "February", `03` = "March", `04` = "April", `05` = "May", `06` = "June", `07` = "July", `08` = "August", `09` = "September", `10` = "October", `11` = "November", `12` = "December")) %>% 
  
  mutate(
    president = gop , dez   ) %>%
  select(-prez_dem, -prez_gop, -day)

#How do I create the president variable with the two options based on the other two variables?
#How do I change month, day, and year to numeric?

view(national_politicians)

s_p = 
  read_csv("data/snp.csv") %>%
  clean_names(dat = .) %>%
  separate(col = date,   into = c("month",      "day", "year"), sep = "/") %>%
  arrange(year, month) %>%
  select(year, month, everything())

#How do I change month, day, and year to numeric?
#How do I add the two digits of year to every value within the year column?

view(s_p)

unemployment =
  read_csv("data/unemployment.csv") %>%
  clean_names(dat = .) %>%
  as.character(unemployment, "Jan", "Feb", "Mar", "Apr", "May", "Jun")
  pivot_longer(
    unemployment, 
    Jan:Dec,
    names_to = "month", 
    values_to = "unemployment_percentage")

#how do I change variables to character?

view(unemployment)
